Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-04-03 thesis/chapters/ch_chapter4.tex
Finished draft (mostly). Just a few spots to compute numbers and check references

@@

\section{Results}

Perhaps the three most important parameters to quantify for each enrichment event are: 1) what fraction of released metals are immediately available for star formation, 2) what fraction of metals are carried out of the galaxy in outflows, and 3) how and over what timescale do metals cool from hot phases into star forming gas. We discuss each of these points in Section~\ref{ch4:sec:ISM CGM}. In Section~\ref{ch4:sec:spreads} we go further and address the relative homogeneity of the metals retained by the ISM for each enrichment source.

@@

For clarity we have focused on the \textit{mean} evolution of individual enrichment events at fixed injection energy in the results thus far. We turn now to discuss how much variety exists among these individual events. For each of the discussed runs (from set \runonenu), Figure~\ref{ch4:fig:CGM_CNM_variance} shows the fraction of metals ejected from the galaxy (left) and the fraction of metals in the ISM within the CNM (center) as a function of source energy at two different times, 20~Myr (blue) and 100~Myr (orange). In the right panel, we show how the spread at each of these times varies for each event. The averaged values at each energy are shown as points, and the bars denote the min/max value for each event energy.\footnote{While the standard deviation or IQR may be more informative here, rather than min / max, the sample size at each energy, with the exception of the E46 runs, is too small to reliably do so.} As shown, there is substantial variation at each energy. This is particularly striking for the E46 events, which seemingly can have effectively any value for $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ at any point during the evolution. Finally, the right panel shows that the degree of inhomogeneous mixing for each event varies significantly as well. However, there is a more clear trend for each of the mean, min, and max values of the spread at 100~Myr (decreasing as a function of increasing $E_{\rm ej}$).

@@

It is clear that the trends discussed thus far applies to the average behavior of events at a fixed $E_{\rm ej}$, but not for an individual event. In addition, a higher $E_{\rm ej}$ seems to increase the minimum $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ expected, but the behavior of any one event must depend on additional factors. But what causes this variation among individual events? Do additional properties about the injection site (ISM density, position, local SFR), determine the subsequent evolution? Or is this the result of a complex, non-deterministic set of complex interactions the metals from a given event experience over their lifetime in the galaxy? We investigate the dependence on a few of these properties below.

\subsection{Dependence on Radial Position}
\label{ch4:sec:radial position}
We place each source in regular, but arbitrary, positions in the galaxy without consideration for the local ISM conditions at injection. Sources are placed at the center of the galaxy ($r = 0$~pc) or at various radii (at cardinal positions at 300~pc, and 600~pc), all in the mid-place of the galaxy. Since the E46 events were likely to not self-interact, we additionally placed events at $r = 100$~pc and one event at each $r$ at about one scale height above the disk ($z = 50$~pc). For each \runonenu~ run, we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ as a function of radial position in the galaxy for each event in Figure~\ref{ch4:fig:radial position}. The points, error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.

 \begin{figure*}
   \centering
   \includegraphics[width=0.95\linewidth]{figures/ch4/II_radial_dependence}
   \caption{The dependence of $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ on the radial position of each event at two different times in the evolution of the galaxy. The meaning of each point, the error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.}
   \label{ch4:fig:radial position}
 \end{figure*}

Across runs, there is no clear dependence on the radial position of the event and the fraction of metals ejected beyond the galaxy for any of the displayed runs at either 20~Myr or 100~Myr. There is a slight increase in the minimum fraction of metals ejected from the galaxy as a function of radius for the higher energy runs ($>10^{50}$~erg). This perhaps not surprising for the largest radii as 600~pc is right near the very edge of the galaxy's disk. But again, the trend is weak and there is clearly still a great variance at both fixed $E_{\rm ej}$ and $r$. The fraction of metals in the ISM that are contained within the CNM (middle row) does not show any obvious trend at E46, but does seem to generally be lowest for the higher energy sources at the outskirts of the galaxy. Finally, the radial position does seem to somewhat affect the mixing behavior. Sources at the outskirts of the galaxy tend to be more inhomogeneous than sources at smaller radii, especially at 20~Myr for the higher energy events, and the variation in behavior for these sources decreases substantially, particularly at 100~Myr. It is unclear what exactly drives these mixing differences. If feedback-driven turbulence is the dominate driver of ISM mixing in this dwarf galaxy -- as it likely is -- then sources near the center are more likely to be affected by more feedback events (due to a slightly higher $\Sigma_{\rm SFR}$) and mix more efficiently.

@@

We compute the average ISM properties within a 4-zone radius ($4 \times dx$=7.2~pc) around each injection site to examine any potential correlations in evolution with ISM properties.\footnote{Note that the injection region for each event is a 3-cell radius sphere mapped onto the grid with a CIC interpolation scheme.} In Figure~\ref{ch4:fig:ISM_variance} we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ for each \runonenu~ run as a function of the average number density ($<n>$) in the injection region within 0.1~Myr (our time resolution) of the event. The E46 events clearly do a better job sampling the ISM density PDF, with a slight trend of decreasing $f_{\rm CGM}$ (more retained metals) with local ISM number density. This difference is even true after 100~Myr of evoltuion, but still with significant variation at fixed $n$. The number statistics limit a conclusive result for the higher energy events, but generally there does not appear to be an obvious trend in behavior of $f_{\rm CGM}$ as a function of $n$. This is perhaps because these energies are sufficient to heat up and dramatically reduce the density of the ISM in the injection region anyway, such that $n$ before the event is not as important as it is for the E46 events which are much less dynamically important. $f_{\rm CNM} / f_{\rm Disk}$ does not display any significant variation with $n$ at 100~Myr, but is slightly elevated at 20~Myr at higher densities. There is still a variety of behavior at fixed $n$. Finally, we note no obvious trends of the spread of the metal PDFs as a function of $n$ for all but the E46 run. Again, the local ISM densities is likely not too dynamically relevant to the higher energy events -- hence their lack of variation with $n$ -- but this is not true for E46. There is a fairly clear trend of increasing inhomogeneity in the CNM as a function of the initial $n$ in the injection region for E46, that is not present at 20~Myr, but is obvious at 100~Myr.

 \begin{figure*}
   \centering
   \includegraphics[width=0.95\linewidth]{figures/ch4/II_n_dependence}
   \caption{The variation of $f_{\rm CGM}$, $f_{\rm CNM} / f_{\rm Disk}$, and the spread of the metal fraction PDFs in the ISM as a function of the local ISM number density ($n$) within the injection region of the event just prior (within 0.1~Myr) of the event. Points, lines, and line colors mean the same as shown in Figure~\ref{ch4:fig:CGM_CNM_variance}}
   \label{ch4:fig:ISM_variance}
 \end{figure*}

@@

\section{Discussion and Conclusions}

@@

Variations with individual events in the above results suggests that it would be challenging to make any conclusive statements about the enrichment behavior of single enrichment sources. \textbf{Go back and look again at one-shot enrichment paper} This is problematic for interpreting the chemical abundances in UFDs that are likely to have been host to just a single, exotic enrichment event. Using these observations to constrain the total nucelosynthetic yields of these sources requires implicit assumptions about how quickly metals are available in star forming gas, how homogenously they are distributed throughout the galaxy, and what fraction are ejected from the galaxy. It is likely that none of these quantities can be determined as simple parameters based upon the feedback properties of a single enrichment source or even the global properties of a galaxy (such as $M_{\rm halo}$, $\Sigma_{\rm SFR}$, $\Sigma_{\rm gas}$). However, this study has shown that these parameters can be characterized in an averaged sense, even though the exact result is subject to substantial stochastic effects. This, combined with some mechanism to provide this stochasticity, can be used in combination with many runs of a simple galactic chemical evolution model to understand the possible outcomes for various XXXXX. Then, the observed properties of an individual galaxy can be mapped to this XXXXXX

>>>>

The elemental abundances of a galaxy over time are sensitive to the nuclear physics and stellar astrophysics that determines which stars make what elements and when. The stellar abundance patters within a given galaxy also depend on the details of how those metals are released into the ISM through various forms of stellar feedback, and the hydrodynamic interactions that ultimately mix those elements into the ISM or eject them from the galaxy in galactic winds. Stellar abundance patterns in old, metal poor systems are the result of enrichment from substantially fewer sources than younger, more metal rich stellar populations. Because of this, studying abundance patterns in these environments allows us to place constraints on the yields of individual enrichment events and the nucleosynthetic sites of each element. The lowest mass dwarf galaxies in the Local Group, UFDs, offer some of the best constraints on these processes. These environments can potentially be used to trace and constrain enrichment from the first stars \citep[e.g.][]{FrebelBromm2012,Ji2015,Ritter2015,Jeon2017,Hartwig2018}. and have been used to place constraints on the variety of possible astrophysical sources of r-process enrichment \citep[e.g.][]{Ji2016a,Ji2016b,Ji2018a,Ji2019,Tsujimoto2017,Ojima2018,Duggan2018}. It is likely that chemical enrichment in more massive, metal rich systems with ongoing star formation like the Milky Way is much more uniform, averaging over the variety of complex hydrodynamics and feedback-driven turbulence interactions that may be driving much of the significant variance in stellar abundances seen at low metallicity. Stochasticity, which has been shown to be important for setting the width of stellar abundance patterns low metallicity environments of the Milky Way \citep[e.g.][]{Cescutti2008,Cescutti2014}, plays an even greater role in the evolution of these low mass dwarf galaxies \citep{Applebaum2018}. This complicates the interpretation of abundance patters in these galaxies, particularly for constraining individual enrichment events.

Recent cosmological zoom simulations have attained sufficient resolution to follow the evolution of individual UFDs (or their progenitors at high redshift) (references) and begin investigating their chemical properties \citep{Jeon2017,Corlies2018,Escala2018}. Additional works have conducted more direct investigations into what drives the enrichment process for individual sources using both hydrodynamics simulations \citep{PanScannapiecoScalo2013,Ritter2015,Safarzadeh2017,Hirai2015,Hirai2017,Emerick2018b,Haynes2019} and semi-analytic models \citep{Beniamani2018,KrumholzTing2018}. In \cite{Emerick2018b} we demonstrate that elements released through AGB winds have both wider abundance distributions in the ISM and lower ejection fractions from the galaxy than elements released in supernovae. It is becoming increasingly clear that enrichment is quite inhomogeneous in low mass dwarf galaxies (refs?), and also that the behavior of metals from any one enrichment event depends on a significant number of complex interactions. In spite of this progress, there is still substantial work to be done in understanding the physical processes that drive both evolution of both the mean and width of stellar abundances in low-mass dwarf galaxies.

In this Chapter we utilize the simulations analyzed in prior chapters to conduct a controlled set of ``mixing experiments'' whereby we restart each simulation with enrichment events placed by-hand in order to more directly investigate the evolution of metals from an individual event. While elemental yields were tracked for each star in our simulation, we lacked the necessary Lagrangian information about the metals once they were released into the ISM to be able to trace the evolution of single enrichment events. We investigate primarily how the feedback ejection energy of individual sources ($E_{\rm ej}$) and global star formation rate at the time of enrichment affects how metals are ejected from the galaxy in galactic winds and metal mixing in the ISM. In Section~\ref{ch4:sec:methods} we briefly outline our methods and discuss the setup of these mixing experiments. In Section~\ref{ch4:sec:results} we discuss the results of these experiments and the role of $E_{\rm ej}$, global SFR, radial position of the enrichment event, and local ISM density around each enrichment sources affects the evolution of its metals, and discuss and conclude in Section~\ref{ch4:sec:discussion conclusion}.

@@

\section{Results}
\label{ch4:sec:results}
Perhaps the three most important parameters to quantify for each enrichment event are: 1) what fraction of released metals are immediately available for star formation, 2) what fraction of metals are carried out of the galaxy in outflows, and 3) how and over what timescale do metals cool from hot phases into star forming gas. We discuss each of these points in Section~\ref{ch4:sec:ISM CGM}. In Section~\ref{ch4:sec:spreads} we go further and address the relative homogeneity of the metals retained by the ISM for each enrichment source.

@@

For clarity we have focused on the \textit{mean} evolution of individual enrichment events at fixed injection energy in the results thus far. We turn now to discuss how much variety exists among these individual events. For each of the discussed runs (from set \runonenu), Figure~\ref{ch4:fig:CGM_CNM_variance} shows the fraction of metals ejected from the galaxy (left) and the fraction of metals in the ISM within the CNM (center) as a function of source energy at two different times, 20~Myr (blue) and 100~Myr (orange). In the right panel, we show how the spread at each of these times varies for each event. The averaged values at each energy are shown as points, and the bars denote the min/max value for each event energy.\footnote{While the standard deviation or IQR may be more informative here, rather than min / max, the sample size at each energy, with the exception of the E46 runs, is too small to reliably do so.} As shown, there is substantial variation at each energy. This is particularly striking for the E46 events, which seemingly can have any value of $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ at any point during the evolution. Finally, the right panel shows that the degree of inhomogeneous mixing for each event varies significantly as well. However, there is a more clear trend for each of the mean, min, and max values of the spread at 100~Myr (decreasing as a function of increasing $E_{\rm ej}$).

@@

It is clear that the trends discussed thus far applies to the average behavior of events at a fixed $E_{\rm ej}$, but not for an individual event. In addition, a higher $E_{\rm ej}$ seems to increase the minimum $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ expected, but the behavior of any one event must depend on additional factors. But what causes this variation among individual events? Do additional properties about the injection site (ISM density, position, local SFR), determine the subsequent evolution? Or is this the result of a non-deterministic set of complex interactions the metals from a given event experience over their lifetime in the galaxy? We investigate the dependence on a few of these properties below.

\subsection{Dependence on Radial Position}
\label{ch4:sec:radial position}
We place each source at regular, but arbitrary, positions in the galaxy without consideration for the local ISM conditions at injection. Sources are placed at the center of the galaxy ($r = 0$~pc) or at various radii (at cardinal positions at 300~pc, and 600~pc), all in the mid-place of the galaxy. Since the E46 events were likely to not self-interact, we additionally placed events at $r = 100$~pc and one event at each $r$ at about one scale height above the disk ($z = 50$~pc). For each \runonenu~ run, we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ as a function of radial position in the galaxy for each event in Figure~\ref{ch4:fig:radial position}. The points, error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.

 \begin{figure*}
   \centering
   \includegraphics[angle=90,width=0.99\linewidth]{figures/ch4/II_radial_dependence}
   \caption{The dependence of $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ on the radial position of each event at two different times in the evolution of the galaxy. The meaning of each point, the error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.}
   \label{ch4:fig:radial position}
 \end{figure*}

Across runs, there is no clear dependence on the radial position of the event and the fraction of metals ejected beyond the galaxy for any of the displayed runs at either 20~Myr or 100~Myr. There is a slight increase in the minimum fraction of metals ejected from the galaxy as a function of radius for the higher energy runs ($>10^{50}$~erg). This is perhaps not surprising for the largest radii as 600~pc is right near the very edge of the galaxy's disk. But again, the trend is weak and there is clearly still a great variance at both fixed $E_{\rm ej}$ and $r$. The fraction of metals in the ISM that are contained within the CNM (middle row) do not show any obvious trend at E46, but does seem to generally be lowest for the higher energy sources at the outskirts of the galaxy. Finally, the radial position does seem to affect the mixing behavior. Sources at the outskirts of the galaxy tend to be more inhomogeneous than sources at inner radii, especially at 20~Myr for the higher energy events, and the variation in behavior for these sources decreases substantially, particularly at 100~Myr. It is unclear what exactly drives these mixing differences. If feedback-driven turbulence is the dominate driver of ISM mixing in this dwarf galaxy -- as it likely is -- then sources near the center are more likely to be affected by more feedback events (due to a slightly higher $\Sigma_{\rm SFR}$) and mix more efficiently.

@@

We compute the average ISM properties within a 4-zone radius ($4 \times dx$~=~7.2~pc) around each injection site to examine any potential correlations in evolution with ISM properties.\footnote{Note that the injection region for each event is a 3-cell radius sphere mapped onto the grid with a CIC interpolation scheme.} In Figure~\ref{ch4:fig:ISM_variance} we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ for each \runonenu~ run as a function of the average number density ($<n>$) in the injection region within 0.1~Myr (our time resolution) of the event. The E46 events clearly do a better job sampling the ISM density PDF, with a slight trend of decreasing $f_{\rm CGM}$ (more retained metals) with local ISM number density. This difference is even true after 100~Myr of evolution, but still with significant variation at fixed $n$. The number statistics limit a conclusive result for the higher energy events, but generally there does not appear to be an obvious trend in behavior of $f_{\rm CGM}$ as a function of $n$. This is perhaps because these energies are sufficient to heat up and dramatically reduce the density of the ISM in the injection region regardless of $n$. This is not the case for the E46 events. $f_{\rm CNM} / f_{\rm Disk}$ does not display any significant variation with $n$ at 100~Myr, but is slightly elevated at 20~Myr at higher densities. There is still a variety of behavior at fixed $n$. Finally, we note no obvious trends of the spread of the metal PDFs as a function of $n$ for all but the E46 run. Again, the local ISM densities is likely not too dynamically relevant to the higher energy events -- hence their lack of variation with $n$ -- but this is not true for E46. There is a fairly clear trend of increasing inhomogeneity in the CNM as a function of the initial $n$ in the injection region for E46, that is not present at 20~Myr, but is obvious at 100~Myr.

 \begin{figure*}
   \centering
   \includegraphics[angle=90,width=0.99\linewidth]{figures/ch4/II_n_dependence}
   \caption{The variation of $f_{\rm CGM}$, $f_{\rm CNM} / f_{\rm Disk}$, and the spread of the metal fraction PDFs in the ISM as a function of the local ISM number density ($n$) within the injection region of the event just prior (within 0.1~Myr) of the event. Points, lines, and line colors mean the same as shown in Figure~\ref{ch4:fig:CGM_CNM_variance}}
   \label{ch4:fig:ISM_variance}
 \end{figure*}

@@

\section{Discussion and Conclusions}
\label{ch4:sec:discussion conclusion}

@@

Variations with individual events in the above results suggests that it would be challenging to make any conclusive statements about the enrichment behavior of single enrichment sources. \textbf{Go back and look again at one-shot enrichment paper} This is problematic for interpreting the chemical abundances in UFDs that are likely to have been host to just a single, exotic enrichment event. Using these observations to constrain the total nucleosynthetic yields of these sources requires implicit assumptions about how quickly metals are available in star forming gas, how homogeneously they are distributed throughout the galaxy, and what fraction are ejected from the galaxy. It is likely that none of these quantities can be determined as simple parameters based upon the feedback properties of a single enrichment source or even the global properties of a galaxy (such as $M_{\rm halo}$, $\Sigma_{\rm SFR}$, $\Sigma_{\rm gas}$). However, this study has shown that these parameters can be characterized in an averaged sense, even though the exact result is subject to substantial stochastic effects. This, combined with some mechanism to provide this stochasticity, can be used in combination with many runs of a simple galactic chemical evolution model to understand the possible outcomes for various galaxies. Developing such a model will be a powerful tool for interpreting the irregular stellar abundance patterns observed in low mass dwarf galaxies in the nearby Universe.
